Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-04-10 doc/manuscript_hkk_drought_sensitivity.Rmd
remove text on neg growth #28

*(See [issue #28](https://github.com/forestgeo/growth-precip-thailand/issues/28))*
To account for potential biases in dendrometer band increments from water-related growth or shrinkage, we also calculated growth occurrence using a zero-growth assumption [@zweifel_are_2016]. 
Timing of measurement and seasonal dynamics in tree hydration and bark dynamics could contribute to negative values in dendrometer band increments, unlike woody growth from tree ring datasets [@chitra-tarak_it_2015]. 
Using a zero-growth model, we calculated growth occurrence as 1 under positive diameter change and any change <=0 as 0. 
To account for biases introduced by timing of measurements and seasonal dynamics, we used this simplifying assumption to corroborate the patterns observed with the sensitivity metric.

@@

Drought years were associated with lower growth across the plot. 
In 2010, the median annual increment was `r round(median(tree.time$inc_annual[tree.time$yr==2010]*10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2010]*10, na.rm=T)-median(tree.time$inc_annual*10, na.rm=T))/median(tree.time$inc_annual*10, na.rm=T)),2)`% lower than the median across all years. 
The drought year 2015 had the lowest median annual increment `r round(median(tree.time$inc_annual[tree.time$yr==2015]*10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2015]*10, na.rm=T)-median(tree.time$inc_annual*10, na.rm=T))/median(tree.time$inc_annual*10, na.rm=T)),2)`% lower than the median across all years. 
While median species responses to the 2010 drought were variable, all species had lower increment than a typical year in 2015. 
Besides decrease in magnitude of growth, the proportion of trees with negative growth was also higher in the drought years (Fig S4). 
Across the years, on average, `r round(100*mean(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% of the trees had negative growth, with a minimum of `r round(100*min(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% in `r names(sort(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))))[1]`. 
The drought year 2015 had the maximum proportion of trees with negative growth in the timeseries, `r round(100*sum(tree.time$inc_annual[tree.time$yr==2015]<0)/nrow(tree.time[tree.time$yr==2015,]),2)`%.
<!-- *(See [issue #28](https://github.com/forestgeo/growth-precip-thailand/issues/28))* -->

>>>>

<!-- *(See [issue #28](https://github.com/forestgeo/growth-precip-thailand/issues/28))* -->
<!-- To account for potential biases in dendrometer band increments from water-related growth or shrinkage, we also calculated growth occurrence using a zero-growth assumption [@zweifel_are_2016]. 
Timing of measurement and seasonal dynamics in tree hydration and bark dynamics could contribute to negative values in dendrometer band increments, unlike woody growth from tree ring datasets [@chitra-tarak_it_2015]. 
Using a zero-growth model, we calculated growth occurrence as 1 under positive diameter change and any change <=0 as 0. 
To account for biases introduced by timing of measurements and seasonal dynamics, we used this simplifying assumption to corroborate the patterns observed with the sensitivity metric.   -->

@@

Drought years were associated with lower growth across the plot. 
In 2010, the median annual increment was `r round(median(tree.time$inc_annual[tree.time$yr==2010]*10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2010]*10, na.rm=T)-median(tree.time$inc_annual*10, na.rm=T))/median(tree.time$inc_annual*10, na.rm=T)),2)`% lower than the median across all years. 
The drought year 2015 had the lowest median annual increment `r round(median(tree.time$inc_annual[tree.time$yr==2015]*10, na.rm=T),2)` mm, `r round(-100*((median(tree.time$inc_annual[tree.time$yr==2015]*10, na.rm=T)-median(tree.time$inc_annual*10, na.rm=T))/median(tree.time$inc_annual*10, na.rm=T)),2)`% lower than the median across all years. 
While median species responses to the 2010 drought were variable, all species had lower increment than a typical year in 2015. 
<!-- Besides decrease in magnitude of growth, the proportion of trees with negative growth was also higher in the drought years (Fig S4).  -->
<!-- Across the years, on average, `r round(100*mean(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% of the trees had negative growth, with a minimum of `r round(100*min(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))),2)`% in `r names(sort(tapply(tree.time$inc_annual, tree.time$yr, function(x) sum(x < 0) / length(x))))[1]`.  -->
<!-- The drought year 2015 had the maximum proportion of trees with negative growth in the timeseries, `r round(100*sum(tree.time$inc_annual[tree.time$yr==2015]<0)/nrow(tree.time[tree.time$yr==2015,]),2)`%. -->
<!-- *(See [issue #28](https://github.com/forestgeo/growth-precip-thailand/issues/28))* -->
